Learning by Teaching a Robot:\\The Case of Handwriting

### 2016-01-12 main.tex
List research questions in intro -- Pierre's suggestion

The following sections of the paper go into the details. We first provide a brief
overview of the robotic system and the interaction it induces. We also present
the machine-learning techniques that allow the robot to learn from the children,
followed by the implementation on a {\sc nao} robot.

We then present and report on the field experiments that we have conducted over
the last two years, including four studies at schools, one longer experiment with
eight children in an occupational therapy clinic, and two one-month
long case studies. While the focus of the school experiments was mostly the
technical validation and data acquisition, the three other experiments involved
children with actual deficits, and gave us initial insights into the relevance and
effectiveness of our approach.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation of the Interaction}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{System Validation at Schools}

@@

The other three experiments (the group study at the clinic and the two
case-studies) provide further qualitative evidence of engagement over longer
interaction periods. In particular, as reported hereafter, the two 4-week long
case studies that have been conducted so far show that our system can sustain
children's engagement over durations (5 hours) that are getting closer to what is
expected to have an impact in a real therapeutic context.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Clinic Study: How Children Take on the Role of a Teacher}\label{normandie}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 1: Vincent}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 2: Thomas}

>>>>

\subsection{Research Questions}

These initial considerations can be turned into a set of research questions.

\paragraph{Technical Feasibility} The technical feasibility of our system (and
its actual implementation) is the first question -- actually a prerequisite --
that we want to address. The robotic system needs to be able to produce
handwriting in a believable manner; the legibility of its handwriting (\ie the
apparent skill of the robot) needs to be easy to parametrise and must evolve
over time in a controlled manner; the robot needs to acquire, recognize and
learn from handwriting demonstrations that may be hardly legible (due to the
impairments of our target children); the design and implementation of the
robot's behaviors must effectively sustain and support a complex and long-term
interaction with children; and finally, in contrast to previous work, we aim at
building a fully autonomous agent that can be deployed and used by non-experts
(teachers, therapists). Taken together, these technical objectives are
ambitious and concern a range of different technical subcomponents.

\paragraph{Acceptance} A second research question that we study here relates to
the acceptance of such a system in educative and medical contexts. Remediation
to cognitive deficits is a sensitive question for the caretakers, all the more
so when working with children. Introducing a robot-based system in such
situations is not self-evident, especially as the mentor-protégé relationship
further relies on building strong affective and cognitive bonds between the
child and the robot (\ie the robot is not simply experienced as a passive tool,
as would be a computer program for instance). By establishing collaborations
with teachers, parents and therapists, we aim at going to the practitioners'
field level and explore the conditions for such a robot-based system to be
accepted.

\paragraph{Engagement} We have already mentioned that handwriting remediation
requires exercising over long periods of time. In terms of human-robot
interaction, this translates into a third research question: how to sustain
several hours of interaction while performing a task (handwriting) which is
essentially a boring routine? This is usually studied in term of
\emph{engagement}: how to engage, and thereafter sustain the engagement, of a
child into an interaction with a robot? As already hinted, we hypothesize that
the \emph{learning by teaching} paradigm may prove an effective lever to get the
child to commit to help the robot over an extended period of time. This needs to
be evidenced, which incidentally calls for actually running interaction
experiments over long durations (several hours spread over several weeks).

\paragraph{Remediation Efficacy} Finally, the last, key research question is: is
our approach effective in helping children to recover from handwriting deficits?
Answering this question in a broad, general way is unlikely as {\it 1)} the
observed deficits vary a lot between the children (no single metric) and have
complex, compound causes, and, {\it 2)} ``learning how to write'' is not a
monolithic, linear skill that could be easily rated from 0 to 1.  However, by
crossing qualitative assessments made by the caretakers with quantitative
measures of visual similarity between the letters drawn by the children and some
reference templates, we aim at assessing the impact of our system on the
children' handwriting, at least in terms of trend.

\vspace{2em}
The following sections explore and attempt to answer these
questions. We first provide an
overview of the robotic system and the interaction it induces
in Section~\ref{impl}. We present there
the machine-learning techniques that allow the robot to learn from the children,
as well as the actual implementation of the system on a {\sc nao} robot.

We then present and report on the field experiments that we have conducted over
the last two years, including four studies at schools
(Section~\ref{studies_school}), one longer experiment with
eight children in an occupational therapy clinic (Section~\ref{study_normandy}), and two one-month
long case studies (Sections~\ref{vincent} and~\ref{thomas}). While the focus of the school experiments was mostly the
technical validation and data acquisition, the three other experiments involved
children with actual deficits, and gave us initial insights into the relevance and
effectiveness of our approach.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation of the Interaction}
\label{impl}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{System Validation at Schools}
\label{studies_school}

@@

The other three experiments (the group study at the clinic and the two
case-studies) provide further qualitative evidence of engagement over longer
interaction periods. In particular, as reported hereafter, the two 4-week long
case studies that have been conducted so far show that our system can sustain
children's engagement over durations (5 hours) that are closer to what is
expected to have an impact in a real therapeutic context.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Clinic Study: How Children Take on the Role of a Teacher}\label{normandie}
\label{study_normandy}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 1: Vincent}
\label{vincent}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 2: Thomas}
\label{thomas}
